|

### 2017-11-23 boettiger-eddelbuettel.Rmd
ok, most other minor issues from 3_2 now addressed.

Docker gives users very convenient access to pre-configured and
pre-built _binary_ images that "just work".  This allows R users to
access a wider-variety of ready-to-use environments than provided by
either the R Project itself or, say, their distribution which will
generally focus on one (current) release.  For example, R users on
Windows may run RStudio Server or Shiny Server locally just by launching
a single command (once Docker itself is installed). Another common
use-case is access to R-devel without affecting the local system.
Here, we detail some of the principal use cases motivating these
containerized versions of R environments, and the design principles that
help make them work.

@@

For instance, on any platform wiht Docker installed, the following Docker
command will launch a Rocker container providing the 
RStudio-server environment over a web interface.

@@

The `docker run` option `-p` sets the port on which RStudio will appear, for which 8787 is
the RStudio default (adding your user to the docker group to avoids the
need for a `sudo` command to call `docker`: `sudo usermod -g docker $USER`). 
Many academic and commercial cloud providers make it
possible to execute such code snippets when a container is launched, without
ever needing to `ssh` into the machine. The user may log into the server merely
by pasting its IP address or DNS name (followed by the chosen port, _e.g._, `:8787`)
into a browser and entering the appropriate password. This provides the user with a familiar,
interactive environment running on a remote machine while requiring
a minimum of expertise.

@@

Having a shared, transparent computational environment created by a publicly hosted, reproducible
recipe facilitates community input into configuration details. R and many of its packages
and related software can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth.  While this flexibility reflects varying needs,
many users rely on default settings which may be optimized more for simplicity of installation
than performance.  The Rocker recipes reflect significant community input on these choices,
as well as the considerable experience and expertise of the R Debian maintainer of 20 years
in what configuration options to use.  This helps create a more finely tuned, optimized reference
implementation of the R environment as well as a platform for comparing and discussing these 
concerns which are often overlooked elsewhere.  Issues and Pull Requests on the Rocker 
repositories on GitHub attest to these discussions and improvements.
Widespread use of the Rocker image helps promote both testing of these choices and 
contributions further tweaking the configuration from many members of the R community.

@@

To make extensions transparent and persistent, Rocker images can be extended by
any user by writing their own Dockerfiles based on an appropriate Rocker image.
The Dockerfiles in the Rocker stack should themselves provide a simple of example
of this.  A user begins by selecting an appropriate base image for their needs: 
if the RStudio\textsuperscript{\textregistered} interface is desired, a user might start with `FROM rocker/rstudio`;
an image for testing a particular C code in an R package might use `FROM rocker/r-devel-san`,
an image for reproducing a data analysis will probably select a stable version tag
in addition to an appropriate base library, _e.g._,: `FROM rocker/tidyverse:3.4.1`
Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. Details on how to extend Rocker images can
be found at <https://rocker-project.org>.

@@

The Rocker project consists of a suite of images built automatically by
and hosted on the Docker Hub, <https://hub.docker.com/r/rocker>.  Source
Dockerfiles, supporting scripts and documentation are hosted on GitHub
under the organization `rocker-org`, <https://github.com/rocker-org>.
The issue tracker and pull requests are used for community input,
discussions and contributions to these images. The Rocker project wiki
provides a place to synthesize community-contributed documentation,
use-cases, and other knowledge about using the Rocker images.

@@

The `debian:testing`-based stack currently includes seven images
actively maintained by the Rocker development team (Table 1).
`r-base` builds on `debian:testing`, and the other six in the stack
each build directly from `r-base`.  The `r-base` image is unique in
that it is designated as the official image for the R language by the
Docker organization itself.  This official image is reviewed and then
built by employees of Docker Inc based on a Dockerfile maintained by
the Rocker team.  Consequently, users should refer to this image in
Docker commands without an organization namespace, _e.g._, `docker run -t
r-base` to access the official image.  All other images in the Rocker
project are not individually reviewed and built by Docker Inc and must
be referenced using the `rocker` namespace,
_e.g._, `docker run -ti rocker/r-devel`.

@@

This stack also includes the images `shiny` and `rstudio:testing` that
provide Shiny server and RStudio\textsuperscript{\textregistered} server IDE from RStudio\textsuperscript{\textregistered} Inc, built on the
`r-base` image.  RStudio\textsuperscript{\textregistered} and Shiny are registered trademarks of RStudio
Inc, and their use and the distribution of their software in binary form on
Docker Hub has been granted to the Rocker project by explicit permission
from RStudio. Users should review RStudio's trademark use policy
(<http://www.rstudio.com/about/trademark/>) and address inquiries about
further distribution or other questions to <permissions@rstudio.com>.
The Rocker project also provides images with RStudio\textsuperscript{\textregistered} server and Shiny
server in the stable versioned stack.

@@

The `tidyverse` image contains all required and suggested dependencies
of the commonly-used `tidyverse` and `devtools` R packages, including
external database libraries (_e.g._, MariaDB and PostgreSQL).  Users should
consult the package Dockerfiles or `installed.packages()` list directly
for a complete list of installed packages.  The `verse` library adds
commonly-used dependencies, notably a large but not comprehensive LaTeX
environment and Java development libraries.  Previously, the Rocker
project provided the image `hadleyverse` which was since divided into
`tidyverse` and `verse` through community input.

>>>>

Docker gives users very convenient access to pre-configured and
pre-built _binary_ images that "just work".  This allows R users to
access a wider-variety of ready-to-use environments than provided by
either the R Project itself or, say, their distribution which will
generally focus on one (current) release.  For example, R users on
Windows may run RStudio\textsuperscript{\textregistered} Server or Shiny\textsuperscript{\textregistered} Server locally just by launching
a single command (once Docker itself is installed). Another common
use-case is access to R-devel without affecting the local system.
Here, we detail some of the principal use cases motivating these
containerized versions of R environments, and the design principles that
help make them work.

@@

For instance, on any platform wiht Docker installed, the following Docker
command will launch a Rocker container providing the 
rstudio-server environment over a web interface.

@@

The `docker run` option `-p` sets the port on which RStudio\textsuperscript{\textregistered} will appear, for which 8787 is
the default (adding your user to the docker group to avoids the
need for a `sudo` command to call `docker`: `sudo usermod -g docker $USER`). 
Many academic and commercial cloud providers make it
possible to execute such code snippets when a container is launched, without
ever needing to `ssh` into the machine. The user may log into the server merely
by pasting its IP address or DNS name (followed by the chosen port, _e.g._, `:8787`)
into a browser and entering the appropriate password. This provides the user with a familiar,
interactive environment running on a remote machine while requiring
a minimum of expertise.

@@

Having a shared, transparent computational environment created by a publicly hosted, reproducible
recipe facilitates community input into configuration details. R and many of its packages
and related software can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth.  While this flexibility reflects varying needs,
many users rely on default settings which are most often are optimized more for 
simplicity of installation rather than than performance.
The Rocker recipes reflect significant community input on these choices.
This helps create a more finely tuned, optimized reference
implementation of the R environment as well as a platform for comparing and discussing these 
concerns which are often overlooked elsewhere.  Issues and Pull Requests on the Rocker 
repositories on GitHub attest to some of these discussions and improvements.
In particular, input from the Docker Inc employees through the official approval process
for the `r-base` image, expertise from the Debian R maintainer and other Debian developers,
and both direct and indirect feedback from the experience and user-generated documentation
from many early adopters in the R community has helped shape and strengthen the project
over the past few years.
Widespread use of the Rocker image helps promote both testing of these choices and 
contributions further tweaking the configuration from many members of the R community.

@@

To make extensions transparent and persistent, Rocker images can be extended by
any user by writing their own Dockerfiles based on an appropriate Rocker image.
The Dockerfiles in the Rocker stack should themselves provide a simple of example
of this, (as described in the following section).  
A user begins by selecting an appropriate base image for their needs: 
if the RStudio\textsuperscript{\textregistered} interface is desired, a user might start with `FROM rocker/rstudio`;
an image for testing an R package with compiled code might use `FROM rocker/r-devel-san`,
an image for reproducing a data analysis will probably select a stable version tag
in addition to an appropriate base library, _e.g._,: `FROM rocker/tidyverse:3.4.1`
Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. Details on how to extend Rocker images can
be found at <https://rocker-project.org>.

@@

The Rocker project consists of a suite of images built automatically by
and hosted on the Docker Hub, <https://hub.docker.com/r/rocker>.  Source
Dockerfiles, supporting scripts and documentation are hosted on GitHub
under the organization `rocker-org`, <https://github.com/rocker-org>.
The issue tracker and pull requests are used for community input,
discussions and contributions to these images. The Rocker project wiki,
<https://github.com/rocker-org/rocker/wiki>,
provides a place to synthesize community-contributed documentation,
use-cases, and other knowledge about using the Rocker images.

@@

The `debian:testing`-based stack currently includes seven images
actively maintained by the Rocker development team (Table 1).
`r-base` builds on `debian:testing`, and the other six in the stack
each build directly from `r-base`.  The `r-base` image is unique in
that it is designated as the official image for the R language by the
Docker organization itself.  This official image is reviewed and then
built by employees of Docker Inc based on a Dockerfile maintained by
the Rocker team.  Consequently, users should refer to this image in
Docker commands without an organization namespace, _e.g._, `docker run -ti r-base` 
to access the official image.  All other images in the Rocker
project are not individually reviewed and built by Docker Inc and must
be referenced using the `rocker` namespace,
_e.g._, `docker run -ti rocker/r-devel`.

@@

This stack also includes the images `shiny` and `rstudio:testing` that
provide Shiny server and RStudio\textsuperscript{\textregistered} server IDE from RStudio\textsuperscript{\textregistered} Inc, built on the
`r-base` image.  RStudio\textsuperscript{\textregistered} and Shiny are registered trademarks of RStudio
Inc, and their use and the distribution of their software in binary form on
Docker Hub has been granted to the Rocker project by explicit permission
from RStudio. Users should review RStudio\textsuperscript{\textregistered}'s trademark use policy
(<http://www.rstudio.com/about/trademark/>) and address inquiries about
further distribution or other questions to <permissions@rstudio.com>.
The Rocker project also provides images with RStudio\textsuperscript{\textregistered} server and Shiny
server in the stable versioned stack.

@@

The `tidyverse` image contains all required and suggested dependencies
of the commonly-used `tidyverse` and `devtools` R packages, including
external database libraries (_e.g._, MariaDB and PostgreSQL).  Users should
consult the package Dockerfiles or `installed.packages()` list directly
for a complete list of installed packages.  The `verse` library adds
commonly-used dependencies, notably a large but not comprehensive LaTeX
environment and Java development libraries.  Previously, the Rocker
project provided the image `hadleyverse` which was since divided into
`tidyverse` and `verse` based on community input.
